Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-03 paper/Response.tex

\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage{url}

\usepackage{color}
\newcommand{\revise}[1]{\textcolor{red}{#1}}

@@

\setlength{\voffset}{-0.0in}
\hoffset=-50pt
\topmargin=0pt
\headheight=0pt
\headsep=0pt
\evensidemargin=72pt % 1in+7pt

\setlength{\parskip}{0.5ex}
\setlength{\parindent}{1em}
\setlength{\floatsep}{1pt}
\setlength{\textfloatsep}{2pt}
\setlength{\intextsep}{2pt}

\begin{document}

October 29, 2010

Response to reviewers:

We thank the reviewers for their careful reading of our paper.  Their
suggestions helped us improve the clarity of what we discuss, and even
sparked some new ideas for improving our algorithms.  Below we list
the Reviewers' comments verbatim.  Our responses are shown in red.

@@

\vspace{0.5 in}

Reviewer \#1: The submitted paper features a description and analysis of a
new MapReduce implementation based on MPI.  The features and limitations
of the implementation are covered with respect to other MapReduce
implementations.  The implementation is applied to several
graph-theoretic algorithms, and the performance of these applications
is compared with applications from an established library (Trilinos).

The introductory and supporting text are, overall, very good.  The
paper places MapReduce in its historical context and motivates its
usage.  The outline of strengths is helpful- important features
include support for out-of-core operations that other libraries (e.g.,
Trilinos) often cannot handle and the ability to use MPI directly in
addition to the MapReduce abstraction.  The paper then provides a good
overview of the composition of operations in a MapReduce application
and a summary of the out-of-core mechanisms available.  The figure
(Figure 1) used to provide a high-level schematic of MapReduce data
flows could be enhanced- for example, it should indicate node-local
and communication intensive operations.

\revise{Figure 1 is actually describing an on-processor rearrangement
of data that each processor performs on its local data.  Thus there is
no inter-processor communication involved.}

@@

All performance results are shown as scaling studies for a range of
problem sizes for the six graph algorithms.  The results show that the
MapReduce applications do not compete well with the reference
implementation (Trilinos), but the authors make clear the limitations
involved in using this high-level library which trades some
performance for usability.  Additionally, the implementation supports
very large data sets which can only be processed using its out-of-core
feature set.  The results are presented well.

Since improved performance for small and mid-range problems is not the
contribution of the paper, it could be improved by stressing other
aspects of the system.  Real-world out-of-core problems could be
surveyed.  Code size and complexity could be considered with usability
as a motivation.  Is the MapReduce-based graph algorithm library
easier to use than the Trilinos library?  Should higher-level user
applications be considered?  Can the MapReduce-based inefficiencies be
isolated more concisely?

\revise{We added some of these comparison points to the final
``Lessons Learned'' section.  In particular we highlight the
relatively small line count of the MapReduce algorithms and the
natural out-of-core ability it provides, as compared for example to
Trilinos.  The comment with respect to MR inefficiencies is addressed
below.}

@@

The paper does not have a concise "Conclusion", only a "Lessons Learned"
section.

\vspace{0.5 in}

Reviewer \#2: This is a nice paper addressing an important current topic
with a good quality open-source implementation.  The overall presentation
is good: the description of algorithms and implementations is clear and the
performance evaluation is detailed. As the authors pointed out, the
performance of the MR-MPI implementations of some graph algorithms falls
behind the matrix-based counterparts so it would be interesting to further
investigate where the performance descrepency comes from and if it's
possible to optimize the algorithm and implementation to bridge the
performance gap.

\revise{We instrumented our MR-MPI and Trilinos implementations of
PageRank to show where time is spent during the PageRank iterations.
These results were added in Table 2.  We further instrumented the two
most expensive steps of the PageRank iterations, as shown in Tables 3
and 4.  These experiments showed that considerable time is spent in
reorganizing data from key-value pairs to key-multivalue pairs.  We
added text in sections 4.1 and 5.2 to describe alternative data
layouts that could reduce the number of times the data needs to be
reorganized to improve performance.}

@@

\vspace{0.5 in}

Reviewer \#3: I appreciated the clear exposition of your map/reduce
framework, its decomposition into a number of independent and composable
stages, and the explanation of how out-of-core computations are handled.

I also liked that you presented the map/reduce implementation of a
number of different graph algorithms. However, I thought the
presentation of these algorithms could be improved in two ways: - The
text tends to just explaining what the code does, not why, nor why
this solves the specific problem (4.2 was the worst culprit here - f I
understood correctly, the vertices get labeled by edge counts that are
used to give the edges a direction that helps identify triangles: if
this is correct a little diagram showing this would help...)  - The
detailed algorithms (Figures 4-9) are a little too imprecise to follow
at times. In particular, when multiple map/reduces are involved the
lack of names for the map/reduce operations and their inputs makes
things confusing. Please be very explicit (after all the code itself
is), and Figure 4 at least actually had names for the map-reduce
objects.

\revise{We agree that our notation in Section 4 was deficient.  I
couldn't understand some of the deteails when I re-read it!  So we
significantly re-worked and standardized the notation used in Section
4 to hopefully describe all the graph algorithms more precisely.  In
particular we noted which operations are performed with which
MapReduce objects and their associated key/value data, so that the
flow of data through each algorithm should be more clear.  A
discussion of the MR-MPI library's ability to allow creation of
multiple MapReduce objects was added to Section 2.}

@@

Your performance is quite low when compared to non map/reduce
implementations of the same algorithm. It would be interesting to know the
breakdown of execution time between map, shuffle and reduce phases to get a
better understanding of where your system spends its time (not necessarily
for every single performance measurement, but some representative or
overall measurements would be interesting).

\revise{Reviewer \#2 made this same suggestion; please see the
response above.}

### 2010-11-04 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@662 3fc257bb-8550-0410-81b3-a5c103da83f7

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT matrices ~\cite{RMAT} are recursively
generated graphs with power-law degree distributions.  They are
commonly used to represent web and social networks.  The user
specifices six parameters that define the graph: the number of
vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
that sum to 1.0 and are discussed below.  The algorithm in
Figure~\ref{fig:rmat} uses a single MapReduce object to generate $M$
unique non-zero entries in a sparse $N \times N$ matrix $A$, where
each entry $A_{ij}$ represents an edge between graph vertices
$(V_i,V_j)$.

@@

$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map: \> Generate $M_{\rm remain}/P$ random edges $(i,j)$ on each processor \\
           \> \> output Key = $(i,j)$, Value = NULL \\
\> Collate \\
\> Reduce: \> Remove duplicate edges \\
           \> \> input Key = $(i,j)$, MultiValue = one or more NULLs \\
           \> \> output Key = $(i,j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$

@@

 \caption{\it MapReduce algorithm for R-MAT graph generation on $P$
 processors, using a single MapReduce object.}

@@

In the {\it map()} operation, each of $P$ processors generates a $1/P$
fraction of the desired edges.  A single random edge ($i,j$) is computed
recursively as follows.  Pick a random quadrant of the $A$ matrix with
relative probabilities $a$, $b$, $c$, and $d$.  Treat the chosen
quadrant as a sub-matrix and select a random quadrant within it, in
the same manner.  Repeat this process $n$ times where $N = 2^n$.  At
the end of the recursion, the final ``quadrant'' is non-zero matrix
element $A_{ij}$.

Though A is very sparse, the {\it map()} typically generates some
small number of duplicate edges.  The {\it collate()} and {\it
reduce()} operations remove the duplicates.  The entire
map-collate-reduce sequence is repeated until the number of resulting
key/value pairs $N_{kv} = M$.  For reasonably sparse graphs this
typically takes only a few iterations.

Note that the degree distribution of vertices in the graph depends on
the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
values is larger than the other three, a skewed distribution results.
Variants of the above algorithm can be used when $N$ is not a
power-of-two, to generate graphs with weighted edges (assign a numeric
value to the $A_{ij}$ edge), graphs without self edges (require $i \ne
j$), or graphs with undirected edges (require $i < j$).  Or the
general R-MAT matrix can be further processed by MapReduce operations
to meet these requirements.  For some algorithms below (e.g., triangle
finding and maximal independent set generation), we post-process R-MAT
matrices to create upper-triangular R-MAT matrices representing
undirected graphs.  For each edge $(i,j)$ with $i>j$, we add edge
$(j,i)$, essentially symmetrizing the matrix; we then remove duplicate
edges and self edges, leaving only edges with $i<j$.

By using more than one MapReduce object, we can improve the
performance of the R-MAT generation algorithm, particularly when
out-of-core operations are needed.  In the enhanced algorithm
(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce
object $E_{new}$ and {\it aggregate()} the edges to processors.  We
then add those edges to the MapReduce object $E_{old}$, containing all
previously generated unique edges; the edges of $E_{old}$ are already
aggregated to processors using the same mapping as $E_{new}$.  We can
then use the {\it compress()} function to reduce duplicates locally on
each processor.  The cost of the algorithm is reduced due to the
smaller number of KV pairs communicated in the {\it aggregate()},
compared to the {\it collate()} operation of the original algorithm,
particularly in later iterations of the algorithm where few edges are
created.

@@

$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $E_{new}$: \> Generate $M_{\rm remain}/P$ random edges $(i,j)$ on each processor \\
           \> \> output Key = $(i,j)$, Value = NULL \\
\> Aggregate $E_{new}$ \\
\> Add $E_{new}$ to $E_{old}$ \\
\> Compress $E_{old}$: Remove duplicate edges \\
            \> \> input Key = $(i,j)$, MultiValue = one or more NULLs \\
            \> \> output Key = $(i,j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$

@@

 \caption{\it Enhanced MapReduce algorithm for R-MAT graph generation
 on $P$ processors, using two MapReduce objects $E_{new}$ and
 $E_{old}$.}

>>>>

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT matrices ~\cite{RMAT} are recursively
generated graphs with power-law degree distributions.  They are
commonly used to represent web and social networks.  The user
specifices six parameters that define the graph: the number of
vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
that sum to 1.0 and are discussed below.  The algorithm in
Figure~\ref{fig:rmat} uses a single MapReduce object to generate $M$
unique non-zero entries in a sparse $N \times N$ matrix $G$, where
each entry $G_{ij}$ represents an edge between graph vertices
$(V_i,V_j)$, with $V_i$ an integer from $1$ to $N$.

@@

Input: \\
\> MapReduce object $G$ = edges of graph, initially empty \\
\\
$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $G$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate $G$ \\
\> Reduce $G$: \> Remove duplicate edges \\
\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$ \\
\\
Output: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\

@@

 \caption{\it MapReduce algorithm for R-MAT graph generation on $P$
 processors, using a single MapReduce object $G$.}

@@

In the {\it map()} operation, each of $P$ processors generates a $1/P$
fraction of the desired edges.  A single random edge $(V_i,V_j)$ is
computed recursively as follows.  Pick a random quadrant of the $G$
matrix with relative probabilities $a$, $b$, $c$, and $d$.  Treat the
chosen quadrant as a sub-matrix and select a random quadrant within
it, in the same manner.  Repeat this process $n$ times where $N =
2^n$.  At the end of the recursion, the final ``quadrant'' is a single
non-zero matrix element $G_{ij}$.

Though $G$ is very sparse, the {\it map()} typically generates some
small number of duplicate edges.  The {\it collate()} and {\it
reduce()} operations remove the duplicates.  The entire
map-collate-reduce sequence is repeated until the number of resulting
key/value (KV) pairs $N_{kv} = M$.  For reasonably sparse graphs this
typically takes only a few iterations.

Note that the degree distribution of vertices in the graph depends on
the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
values is larger than the other three, a skewed distribution results.
Variants of the above algorithm can be used when $N$ is not a
power-of-two, to generate graphs with weighted edges (assign a numeric
value to the $G_{ij}$ edge), graphs without self edges (require $i \ne
j$), or graphs with undirected edges (require $i < j$).  Or the
general R-MAT matrix can be further processed by MapReduce operations
to meet these requirements.  For some algorithms below (e.g., triangle
finding and maximal independent set generation), we post-process R-MAT
matrices to create upper-triangular R-MAT matrices representing
undirected graphs.  For each edge $(V_i,V_j)$ with $i > j$, we add
edge $(V_j,V_i)$, essentially symmetrizing the matrix; we then remove
duplicate edges and self edges, leaving only edges with $i < j$.

By using an additional MapReduce object, we can improve the
performance of the R-MAT generation algorithm, particularly when
out-of-core operations are needed.  In the enhanced algorithm
(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce
object $G_{new}$ and {\it aggregate()} the edges to processors.  We
then add those edges to the MapReduce object $G$, containing all
previously generated unique edges; the edges of $G$ are already
aggregated to processors using the same mapping as $G_{new}$.  We can
then use the {\it compress()} operation to reduce duplicates locally
on each processor.  The cost of the enhanced algorithm is less due to
the smaller number of KV pairs communicated in the {\it aggregate()},
compared to the {\it collate()} operation of the original algorithm,
particularly in later iterations of the algorithm where few new edges
are generated.

@@

Input: \\
\> MapReduce object $G$ = edges of graph, initially empty \\
\> MapReduce object $G_{new}$ = new edges of graph, initially empty \\
\\
$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $G_{new}$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> Aggregate $G_{new}$ \\
\> Add $G_{new}$ to $G$, empty $G_{new}$ \\
\> Compress $G$: Remove duplicate edges \\
\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$ \\
\\
Output: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\

@@

 \caption{\it Enhanced MapReduce algorithm for R-MAT graph generation
 on $P$ processors, using two MapReduce objects $G$ and $G_{new}$.}
